---
title: "Milestone Report"
author: "Scott Brenstuhl"
date: "March 12, 2016"
output: html_document
---

```{r, echo=FALSE}
library(tm)
library(readr)
library(stringi)
```

```{r, echo=FALSE}
en_twit <- readLines('final/en_US/en_US.twitter.txt', col_names = FALSE)
en_blog <- readLines('final/en_US/en_US.blogs.txt', col_names = FALSE)
en_news <- readLines('final/en_US/en_US.news.txt', col_names = FALSE)

twit_sample <- readLines('sample/twitter_sample.txt')
```
Goals:
1. Demonstrate that you've downloaded the data and have successfully loaded it in.
2. Create a basic report of summary statistics about the data sets.
3. Report any interesting findings that you amassed so far.
4. Get feedback on your plans for creating a prediction algorithm and Shiny app.

Notes:
This document should be concise and explain only the major features of the data you have identified and briefly summarize your plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager.

You should make use of tables and plots to illustrate important summaries of the data set.

Review Criteria:
Does the link lead to an HTML page describing the exploratory analysis of the training data set?
Has the data scientist done basic summaries of the three files? Word counts, line counts and basic data tables?
Has the data scientist made basic plots, such as histograms to illustrate features of the data?
Was the report written in a brief, concise style, in a way that a non-data scientist manager could appreciate?

```{r}
# Return summary stats
text.stats <- function(text){
    rows.count = length(text)
    words.count = sum(stri_count_words(text))
    paste('Rows:', rows.count, "Words:", words.count)
}

text.stats(en_twit)
text.stats(en_blog)
text.stats(en_news)
```

```{r}
#Remove Twitter handles, hashtags, and RT
twit_sample <- gsub('(@\\S*)', ' ', twit_sample)
twit_sample <- gsub('(#\\S*)', ' ', twit_sample)
twit_sample <- gsub('\\bRT\\b', ' ', twit_sample)
twit_sample <- gsub(':', ' ', twit_sample)

dirty.words.link <- 'https://raw.githubusercontent.com/shutterstock/List-of-Dirty-Naughty-Obscene-and-Otherwise-Bad-Words/master/en'
dirty.words <- readLines(dirty.words.link)
dirty.words <- paste0("\\b(",paste0(dirty.words,collapse="|"), ")\\b") 

text.cleaner <- function(text.body){
    # Gets rid of all the dirty words
    clean.text = gsub(dirty.words, ' ', text.body)
    
    # This section is mostly aimed at twitter but will clean up
    # references to unuseful twitter things in other sources.
    clean.text = gsub('(@\\S*)', ' ', clean.text)
    clean.text = gsub('(#\\S*)', ' ', clean.text)
    clean.text = gsub('\\bRT\\b', ' ', clean.text)
    clean.text = gsub(':', ' ', clean.text)
}

# Return summary stats
text.stats <- function(text.body){
    rows.count = length(text.body)
    words.count = sum(stri_count_words(text.body))
    paste('Rows:', rows.count, "Words:", words.count)
}

text.stats(twit_sample)
```

most popular bi and tri grams by source
